\textbf{ Accurate spectrophotometric estimates of luminosity and distance\\ for luminous red-giant stars }

### 2018-07-29 paper/sp.tex
many words in the data section

% project macros
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\wise}{\project{\acronym{WISE}}}
\newcommand{\zmass}{\project{\acronym{2MASS}}}
\newcommand{\sdssv}{\project{\acronym{SDSS-V}}}

@@

We take the \apogee\ data from what subset?

We cut on $\logg$ how and why? Comment on the trustworthiness of the $\logg$ values.

HOGG: We take the \gaia\ data and apply the parallax offset.

We match \apogee\ to \gaia\ how?

We match the catalog to \zmass\ and \wise\ how?

Figures that show off the data.

We construct the feature vectors $x_n$ how and what is their dimensionality $D$?

>>>>

% text macros
\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
\newcommand{\foreign}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\wise}{\project{\acronym{WISE}}}
\newcommand{\zmass}{\project{\acronym{2MASS}}}
\newcommand{\sdssv}{\project{\acronym{SDSS-V}}}

@@

ACE: We take the \apogee\ data from what subset?
We cut on $\logg$ how and why? Comment on the trustworthiness of the $\logg$ values.

ACE: We match the \apogee\ stars making the $\logg$ cut also to
the \zmass\ ACE WHICH CATALOG by ACE WHAT CRITERION?
This gives, for each matching star $n$,
near-infrared photometry $J_n$, $H_n$, and $K_n$.
In detail we use from the catalog the entries ACE WHAT SPECIFIC CATALOG ENTRIES.

ACE: We match the \apogee+\zmass\ stars to he \wise\ ACE WHICH CATALOG by ACE WHAT CRITERION?
This gives, fore each matching star $n$,
mid-infrared photometry $W_{1n}$ and $W_{2n}$ at ACE WHAT WAVELENGTHS?
In detail we use the ACE WHAT SPECIFIC CATALOG ENTRIES?

We match this full-match catalog to the \gaia\ \acronym{DR2} (HOGG CITE) by what criteria?

For each star $n$, we take from \gaia\ DR2 we the parallax $\varpi_n$
and its uncertainty $\sigma_{\varpi n}$, which will be used in
constructing the inverse-root-luminosity $Q_n$ and associated
uncertainty $\sigma_n$.
Importantly, to each \gaia\ parallax measurement we add a positive
offset of $0.029$\,mas to adjust for the under-estimation of
parallaxes reported by the \gaia\ Collaboration (HOGG CITE LINDEGREN).
We also take from the \gaia\ data the photometric data $G_n$
(\code{phot_g_mean_mag}) and $G_{Bn}-G_{Rn}$ (\code{bp_rp}), which will
become part of the feature vector $x_n$.
We do not use the entire \gaia\ \acronym{DR2} Catalog but instead the subset
that meets ACE WHAT ADDITIONAL PHOTOMETRIC CUTS?

We need complete feature-vector information for all stars.  For this
reason, we define the Parent Sample to be the set of all stars that
meet the \apogee\ and \gaia\ cuts and also
have the complete set of photometry: $G_n$, $G_{Bn}-G_{Rn}$, $J_n$,
$H_n$, $K_n$, $W_{1n}$, and $W_{2n}$ and also an apogee spectrum.
This Parent Sample contains ACE HOW MANY STARS? and is shown in HOGG WHAT FIGURES?

Every Parent Sample object gets, in addition, a randomly assigned binary
label (A or B).
This is used for two-fold validation and jackknife.
In short, we will use the model trained on the A data to assign luminosities
and distances to the B data and \foreign{vice versa}.
The Parent Sample A has HOGG entries and Parent Sample B has HOGG entries.

For training, we need A and B Training Sets.
We define the Training Set stars to be Parent Sample stars that also
have a measured \gaia\ parallax.
The Training-Set stars, in addition to meeting all Parent-Sample cuts,
also must meet ACE WHAT ADDITIONAL CUTS to ensure that the parallax
measurements are good.
The training sets are shown in HOGG WHAT FIGURES?

But---as we have emphasized above---we do not cut ever on parallax or
parallax error or their ratio. For this reason, most of the training set
do not have significantly measured parallaxes and HOGG percent even have
negative parallaxes!
But we will perform our training such that it will be unbiased in these
circumstances.
Indeed it would be necessarily biased if ever we did cut on parallax or
parallax signal-to-noise.

For every star $n$ in the full Parent Sample we construct the feature
vector $x_n$ as HOGG PUT EQUATION HERE.
The feature vectors live in a $D$-dimensional space where $D=$HOGG.
For every star $n$ in either of the Training Sets we also construct a
regression datum $Q_n$ and uncertainty $\sigma_n$ according to HOGG WHAT EQUATION ABOVE?
